---
title: "1mimic_derive"
author: "Andrew Hanson"
output: html_document
---

```{r setup, include=FALSE}
# run this program
# Rscript -e "rmarkdown::render('1mimic_derive.Rmd')"

knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(dbplyr)

row_limit <- -1
```

## Admissions

```{r read services}
`physionet-data.mimic_hosp.services` <- read.csv(file='../../hosp/services.csv.gz', nrows=row_limit)
```

```{r read_admit_data}
`physionet-data.mimic_icu.icustays` <- read.csv(file='../../icu/icustays.csv.gz', nrows=row_limit)

# list of hospitalizations with ICU stays - keep only these patients and admissions
icustays_lst <- `physionet-data.mimic_icu.icustays` %>%
  dplyr::select(subject_id, hadm_id) %>%
  distinct()

`physionet-data.mimic_core.patients` <- read.csv(file='../../hosp/patients.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, 'subject_id')

`physionet-data.mimic_core.admissions` <- read.csv(file='../../hosp/admissions.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))
```

```{r}
icustays_db <- tbl_memdb(`physionet-data.mimic_icu.icustays`)
patients_db <- tbl_memdb(`physionet-data.mimic_core.patients`)
admissions_db <- tbl_memdb(`physionet-data.mimic_core.admissions`)
services_db <- tbl_memdb(`physionet-data.mimic_hosp.services`)

con <- src_memdb()$con
```

```{sql get_surgical_admits, connection=con, output.var="surgical_admits"}
WITH surgflag AS (
    SELECT DISTINCT 
        adm.subject_id
        , adm.hadm_id
        , CASE
            WHEN LOWER(curr_service) LIKE '%surg%' THEN 1 ELSE 0
        END AS surgical
        , ROW_NUMBER() OVER
        (
            PARTITION BY adm.hadm_id
            ORDER BY transfertime
        ) AS serviceorder
    FROM `physionet-data.mimic_core.admissions` adm
    LEFT JOIN `physionet-data.mimic_hosp.services` se
        ON adm.hadm_id = se.hadm_id
)

SELECT 
    s.subject_id
    , s.hadm_id
    , s.surgical
FROM surgflag s
WHERE serviceorder = 1 AND surgical = 1
;
```

## Suspicion of infection

```{r read_antibiotic}
`physionet-data.mimic_hosp.prescriptions` <- read.csv(file='../../hosp/prescriptions.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))

prescriptions_db <- tbl_memdb(`physionet-data.mimic_hosp.prescriptions`) 
```

```{r read_microbiology}
`physionet-data.mimic_hosp.microbiologyevents` <- read.csv(file='../../hosp/microbiologyevents.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))

microbiologyevents_db <- tbl_memdb(`physionet-data.mimic_hosp.microbiologyevents`)
```

### Antibiotics

```{sql derive_antibiotic, connection=con, code=readLines("antibiotic.sql"), output.var="physionet-data.mimic_derived.antibiotic"}

```

```{r}
antibiotic_db <- tbl_memdb(`physionet-data.mimic_derived.antibiotic`)
```

### Suspicion

```{sql derive_suspicion_of_infection, connection=con, code=readLines("suspicion_of_infection.sql"), output.var="physionet-data.mimic_derived.suspicion_of_infection"}

```

```{r}
suspicion_of_infection_db <- tbl_memdb(`physionet-data.mimic_derived.suspicion_of_infection`)

# list of hospitalizations with suspicion of infection - keep only these hospitalizations
susp_lst <- `physionet-data.mimic_derived.suspicion_of_infection` %>%
  filter(suspected_infection == 1) %>% 
  dplyr::select(subject_id, hadm_id) %>% 
  distinct()
```

## Admissions

```{r read_chartevents}
icustays_lst <- `physionet-data.mimic_icu.icustays` %>%
  dplyr::select(subject_id, hadm_id) %>%
  inner_join(susp_lst, c('subject_id','hadm_id')) %>%
  distinct()

mimic_icu.chartevents <- read.csv(file='../../icu/chartevents.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))

# only care about the first 24 hours
mimic_icu.chartevents <- mimic_icu.chartevents %>%
  inner_join(distinct(dplyr::select(`physionet-data.mimic_icu.icustays`, stay_id, intime)), 'stay_id') %>%
  filter(as.POSIXct(intime) - 24*3600 <= as.POSIXct(charttime) &
            as.POSIXct(charttime) <= as.POSIXct(intime) + 36*3600) %>%
  dplyr::select(
    -intime
  )
```

```{r}
chartevents_db <- tbl_memdb(mimic_icu.chartevents)
```

## Urine output

```{sql derive_weight_durations, connection=con, code=readLines("weight_durations.sql"), output.var="physionet-data.mimic_derived.weight_durations"}

```

```{r}
weight_durations_db <- tbl_memdb(`physionet-data.mimic_derived.weight_durations`)
```

```{r read_outputevents}
`physionet-data.mimic_icu.outputevents` <- read.csv(file='../../icu/outputevents.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))

outputevents_db <- tbl_memdb(`physionet-data.mimic_icu.outputevents`)
```

```{sql derive_urine_output, connection=con, code=readLines("urine_output.sql"), output.var="physionet-data.mimic_derived.urine_output"}

```

```{r}
urine_output_db <- tbl_memdb(`physionet-data.mimic_derived.urine_output`)
```

```{sql derive_urine_output_rate, connection=con, code=readLines("urine_output_rate.sql"), output.var="physionet-data.mimic_derived.urine_output_rate"}

```

```{r}
urine_output_rate_db <- tbl_memdb(`physionet-data.mimic_derived.urine_output_rate`)
```

# ICU stay times

```{sql derive_icustay_times, connection=con, code=readLines("icustay_times.sql"), output.var="physionet-data.mimic_derived.icustay_times"}

```

```{r}
icustay_times_db <- tbl_memdb(`physionet-data.mimic_derived.icustay_times`)
```

## Vasopressors

```{r read_inputevents}
`physionet-data.mimic_icu.inputevents` <- read.csv(file='../../icu/inputevents.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))

inputevents_db <- tbl_memdb(`physionet-data.mimic_icu.inputevents`)
```

### Dobutamine

```{sql derive_dobutamine, connection=con, code=readLines("dobutamine.sql"), output.var="physionet-data.mimic_derived.dobutamine"}

```

### Dopamine

```{sql derive_dopamine, connection=con, code=readLines("dopamine.sql"), output.var="physionet-data.mimic_derived.dopamine"}

```

### Epinephrine

```{sql derive_epinephrine, connection=con, code=readLines("epinephrine.sql"), output.var="physionet-data.mimic_derived.epinephrine"}

```

### Milrinone

```{sql derive_milrinone, connection=con, code=readLines("milrinone.sql"), output.var="physionet-data.mimic_derived.milrinone"}

```

### Norepinephrine

```{sql derive_norepinephrine, connection=con, code=readLines("norepinephrine.sql"), output.var="physionet-data.mimic_derived.norepinephrine"}

```

### Phenylephrine

```{sql derive_phenylephrine, connection=con, code=readLines("phenylephrine.sql"), output.var="physionet-data.mimic_derived.phenylephrine"}

```

### Vasopressin

```{sql derive_vasopressin, connection=con, code=readLines("vasopressin.sql"), output.var="physionet-data.mimic_derived.vasopressin"}

```

### Vasoactive agents

```{r}
dobutamine_db <- tbl_memdb(`physionet-data.mimic_derived.dobutamine`)
dopamine_db <- tbl_memdb(`physionet-data.mimic_derived.dopamine`)
epinephrine_db <- tbl_memdb(`physionet-data.mimic_derived.epinephrine`)
milrinone_db <- tbl_memdb(`physionet-data.mimic_derived.milrinone`)
norepinephrine_db <- tbl_memdb(`physionet-data.mimic_derived.norepinephrine`)
phenylephrine_db <- tbl_memdb(`physionet-data.mimic_derived.phenylephrine`)
vasopressin_db <- tbl_memdb(`physionet-data.mimic_derived.vasopressin`)
```

```{sql derive_vasoactive_agent, connection=con, code=readLines("vasoactive_agent.sql"), output.var="physionet-data.mimic_derived.vasoactive_agent"}

```

## Ventilator setting

```{sql derive_ventilator_setting, connection=con, code=readLines("ventilator_setting.sql"), output.var="physionet-data.mimic_derived.ventilator_setting"}

```

```{r}
ventilator_setting_db <- tbl_memdb(`physionet-data.mimic_derived.ventilator_setting`)
```

## Oxygen delivery

```{sql derive_oxygen_delivery, connection=con, code=readLines("oxygen_delivery.sql"), output.var="physionet-data.mimic_derived.oxygen_delivery"}

```

```{r}
oxygen_delivery_db <- tbl_memdb(`physionet-data.mimic_derived.oxygen_delivery`)
```

## Ventilation

```{sql derive_ventilation, connection=con, code=readLines("ventilation.sql"), output.var="physionet-data.mimic_derived.ventilation"}

```

```{r}
ventilation_db <- tbl_memdb(`physionet-data.mimic_derived.ventilation`)
```

## Height, weight, BMI

```{sql derive_height, connection=con, code=readLines("height.sql"), output.var="physionet-data.mimic_derived.height"}

```

```{sql derive_weight, connection=con, code=readLines("weight.sql"), output.var="physionet-data.mimic_derived.weight"}

```

## Age

```{sql derive_age, connection=con, code=readLines("age.sql"), output.var="physionet-data.mimic_derived.age"}

```

```{r join_age}
age_db <- tbl_memdb(`physionet-data.mimic_derived.age`)
```

## Comorbidities

```{r read_diagnoses_icd}
`physionet-data.mimic_hosp.diagnoses_icd` <- read.csv(file='../../hosp/diagnoses_icd.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))

dignoses_db <- tbl_memdb(`physionet-data.mimic_hosp.diagnoses_icd`)
```

### Charlson

```{sql derive_charlson, connection=con, code=readLines("charlson.sql"), output.var="physionet-data.mimic_derived.charlson"}

```

### Other comorbidities

```{sql derive_comorbidities, connection=con, code=readLines("comorbidities.sql"), output.var="physionet-data.mimic_derived.comorbidities"}

```

## Vital signs

```{sql derive_vitalsign, connection=con, code=readLines("vitalsign.sql"), output.var="physionet-data.mimic_derived.vitalsign"}

```

```{r}
vitalsign_db <- tbl_memdb(`physionet-data.mimic_derived.vitalsign`)
```

## Coagulation

```{r read_labevents}
mimic_hosp.labevents <- read.csv(file='../../hosp/labevents.csv.gz', nrows=row_limit) %>%
  inner_join(icustays_lst, c('subject_id','hadm_id'))

labevents_db <- tbl_memdb(mimic_hosp.labevents)
```

```{sql derive_blood_differential, connection=con, code=readLines("blood_differential.sql"), output.var="physionet-data.mimic_derived.blood_differential"}

```

```{r}
blood_differential_db <- tbl_memdb(`physionet-data.mimic_derived.blood_differential`)
```

```{sql derive_coagulation, connection=con, code=readLines("coagulation.sql"), output.var="physionet-data.mimic_derived.coagulation"}

```

```{r}
coagulation_db <- tbl_memdb(`physionet-data.mimic_derived.coagulation`)
```

## Chemistry

```{sql derive_chemistry, connection=con, code=readLines("chemistry.sql"), output.var="physionet-data.mimic_derived.chemistry"}

```

```{r}
chemistry_db <- tbl_memdb(`physionet-data.mimic_derived.chemistry`)
```

## Complete blood count

```{sql derive_cbc, connection=con, code=readLines("complete_blood_count.sql"), output.var="physionet-data.mimic_derived.complete_blood_count"}

```

```{r}
complete_blood_count_db <- tbl_memdb(`physionet-data.mimic_derived.complete_blood_count`)
```

## Blood gasses

```{sql derive_bg, connection=con, code=readLines("bg.sql"), output.var="physionet-data.mimic_derived.bg"}

```

```{r}
bg_db <- tbl_memdb(`physionet-data.mimic_derived.bg`)
```

## Enzyme

```{sql derive_enzyme, connection=con, code=readLines("enzyme.sql"), output.var="physionet-data.mimic_derived.enzyme"}

```

```{r}
enzyme_db <- tbl_memdb(`physionet-data.mimic_derived.enzyme`)
```

## Glasgow coma scale

```{sql derive_gcs, connection=con, code=readLines("gcs.sql"), output.var="physionet-data.mimic_derived.gcs"}

```

```{r}
gcs_db <- tbl_memdb(`physionet-data.mimic_derived.gcs`)
```

## ICU stay times

```{r dplyr_icustay_hourly}
`physionet-data.mimic_derived.icustay_hourly` <- `physionet-data.mimic_derived.icustay_times` %>%
  mutate(
    start_dttm = as.POSIXct(
      as.numeric(as.POSIXct(intime_hr)) - as.numeric(format(as.POSIXct(intime_hr), format="%M"))*60 - as.numeric(format(as.POSIXct(intime_hr), format="%S")), origin='1970-01-01'
    )
  ) %>%
  rowwise() %>%
  mutate(zendtime = list(start_dttm - 24*3600 + seq(0, 48*3600, by = 3600))) %>%
  unnest(zendtime) %>%
  ungroup() %>% 
  mutate(
    hr = as.numeric(difftime(zendtime, intime_hr, units="hours")),
    endtime = as.character(zendtime)
  ) %>%
  dplyr::select(
    stay_id, hr, endtime
  )
```

```{r}
icustay_hourly_db <- tbl_memdb(`physionet-data.mimic_derived.icustay_hourly`)
```

## SOFA

```{sql derive_sofa, connection=con, code=readLines("sofa.sql"), output.var="physionet-data.mimic_derived.sofa"}

```

```{r}
sofa_db <- tbl_memdb(`physionet-data.mimic_derived.sofa`)
```

## SEPSIS 3

```{sql derive_sepsis3, connection=con, code=readLines("sepsis3.sql"), output.var="physionet-data.mimic_derived.sepsis3"}

```

```{r}
sepsis3_db <- tbl_memdb(`physionet-data.mimic_derived.sepsis3`)
```

## Surgical admission and Transplant

```{r read_drg_data}
`physionet-data.mimic_hosp.drgcodes` <- read.csv(file='../../hosp/drgcodes.csv.gz', nrows=row_limit)
```

```{r get_transplants}
transplant_admits <- `physionet-data.mimic_hosp.drgcodes` %>%
  filter(
    grepl('TRANSPLANT', description)
  ) %>%
  mutate(
    transplant = 1
  ) %>%
  dplyr::select(
    subject_id, hadm_id, transplant
  ) %>%
  distinct()
```

### APS

#### First day urine output

```{sql derive_first_day_urine_output, connection=con, code=readLines("first_day_urine_output.sql"), output.var="physionet-data.mimic_derived.first_day_urine_output"}

```

```{r}
first_day_urine_output_db <- tbl_memdb(`physionet-data.mimic_derived.first_day_urine_output`)
```

#### First day labs

```{sql derive_first_day_lab, connection=con, code=readLines("first_day_lab.sql"), output.var="physionet-data.mimic_derived.first_day_lab"}

```

```{r}
first_day_lab_db <- tbl_memdb(`physionet-data.mimic_derived.first_day_lab`)
```

#### First day gcs

```{sql derive_first_day_gcs, connection=con, code=readLines("first_day_gcs.sql"), output.var="physionet-data.mimic_derived.first_day_gcs"}

```

```{r}
first_day_gcs_db <- tbl_memdb(`physionet-data.mimic_derived.first_day_gcs`)
```

#### First day vitalsign

```{sql derive_first_day_vitalsign, connection=con, code=readLines("first_day_vitalsign.sql"), output.var="physionet-data.mimic_derived.first_day_vitalsign"}

```

```{r}
first_day_vitalsign_db <- tbl_memdb(`physionet-data.mimic_derived.first_day_vitalsign`)
```

#### APSIII

```{sql derive_apsiii, connection=con, code=readLines("apsiii.sql"), output.var="physionet-data.mimic_derived.apsiii"}

```

#### SAPSII

```{sql derive_sapsii, connection=con, code=readLines("sapsii.sql"), output.var="physionet-data.mimic_derived.sapsii"}

```

# Create cohort data

```{r create_cohort_data}
# all hospital admissions
out_tb <- `physionet-data.mimic_core.admissions` %>%
  dplyr::select(
    subject_id, hadm_id
  ); dim(out_tb)

# with an ICU admission
out_tb <- out_tb %>%
  inner_join(dplyr::select(
    `physionet-data.mimic_icu.icustays`, subject_id, hadm_id, stay_id), c('subject_id', 'hadm_id')
  ); dim(out_tb)

# with suspicion of infection at any time in the hospitalization
out_tb <- out_tb %>%
  inner_join(icustays_lst, c('subject_id','hadm_id')); dim(out_tb)

# meeting sepsis 3 criteria at any time in the icu admission
tmp <- `physionet-data.mimic_derived.sepsis3` %>%
  filter(sepsis3 == 1) %>%
  dplyr::select(subject_id, stay_id, sepsis3) %>%
  distinct()

out_tb <- out_tb %>%
  inner_join(tmp, c('subject_id','stay_id')); dim(out_tb)
```

```{r}
# join ICU admit related_data
out_tb <- out_tb %>%
  inner_join(`physionet-data.mimic_icu.icustays`, c('subject_id','hadm_id','stay_id')); dim(out_tb)

# join hospital related data
out_tb <- out_tb %>%
  inner_join(`physionet-data.mimic_core.admissions`, c('subject_id','hadm_id')); dim(out_tb)

# sex
tmp <- `physionet-data.mimic_core.patients` %>%
  filter(!is.na(gender)) %>%
  group_by(subject_id) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  dplyr::select(
    subject_id, gender
  )

out_tb <- out_tb %>%
  left_join(tmp, 'subject_id'); dim(out_tb)
```

```{r}
# height
tmp <- `physionet-data.mimic_derived.height` %>%
  filter(!is.na(height)) %>%
  arrange(subject_id, charttime) %>%
  group_by(subject_id) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  dplyr::select(
    subject_id, height
  )

out_tb <- out_tb %>%
  left_join(tmp, 'subject_id'); dim(out_tb)
```

```{r}
# weight
tmp <- `physionet-data.mimic_derived.weight` %>%
  filter(!is.na(weight)) %>%
  arrange(subject_id, stay_id, charttime) %>%
  group_by(subject_id, stay_id) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  dplyr::select(
    stay_id, subject_id, weight
  )

out_tb <- out_tb %>%
  left_join(tmp, c('stay_id','subject_id')); dim(out_tb)
```

```{r}
# age
tmp <- `physionet-data.mimic_derived.age` %>%
  filter(!is.na(age)) %>%
  arrange(hadm_id, subject_id, age) %>%
  group_by(hadm_id, subject_id) %>%
  slice_tail(n=1) %>%
  ungroup() %>%
  dplyr::select(
    hadm_id, subject_id, age
  )

out_tb <- out_tb %>%
  left_join(tmp, c('hadm_id','subject_id')); dim(out_tb)
```

```{r}
# Charlson
tmp <- `physionet-data.mimic_derived.charlson` %>%
  group_by(hadm_id, subject_id) %>%
  slice_head(n=1) %>%
  ungroup()

out_tb <- out_tb %>%
  left_join(tmp, c('hadm_id','subject_id')); dim(out_tb)
```

```{r}
# Comorbidities
tmp <- `physionet-data.mimic_derived.comorbidities` %>%
  group_by(hadm_id, subject_id) %>%
  slice_head(n=1) %>%
  ungroup()

out_tb <- out_tb %>%
  left_join(tmp, c('hadm_id','subject_id')); dim(out_tb)
```

```{r mimic_start_stop_times}
tmp <- `physionet-data.mimic_derived.icustay_times` %>%
  filter(
    intime_hr != ''
  )

out_tb <- out_tb %>%
  inner_join(tmp, join_by(subject_id, hadm_id, stay_id))

out_tb <- out_tb %>%
  mutate(
    intime = intime_hr,
    icu_days = round(as.numeric(difftime(outtime, intime, units='days')), digits=3)
  ) %>%
  dplyr::select(
    -intime_hr, -outtime_hr
  ); dim(out_tb)

out_tb <- out_tb %>%
  distinct()

out_tb <- out_tb %>%
  filter(
    icu_days > 1.85/24
  ); dim(out_tb)
```

```{r mimic_suspicion_of_infection}
tmp <- `physionet-data.mimic_derived.suspicion_of_infection` %>%                           
  filter(
    !is.na(stay_id) & suspected_infection == 1 & suspected_infection_time != ''
  )

tmp <- tmp %>%
  arrange(subject_id, hadm_id, stay_id, suspected_infection_time, antibiotic_time) %>%
  group_by(subject_id, hadm_id, stay_id) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  dplyr::select(
    subject_id, hadm_id, stay_id, suspected_infection_time, antibiotic_time
  )

out_tb <- out_tb %>%
  left_join(tmp, join_by(subject_id, hadm_id, stay_id))

out_tb <- out_tb %>%
  mutate(
    susp = if_else(as.POSIXct(admittime) - 24*3600 <= suspected_infection_time & suspected_infection_time <= as.POSIXct(intime) + 24*3600 &
                     suspected_infection_time <= outtime, 1, 0)
    ); dim(out_tb)

out_tb <- out_tb %>%
  filter(susp == 1); dim(out_tb)
```

```{r}
# surgical
out_tb <- out_tb %>%
  left_join(surgical_admits, c('subject_id','hadm_id')) %>%
  replace_na(list(surgical = 0)); dim(out_tb)

# transplant
out_tb <- out_tb %>%
  left_join(transplant_admits, c('subject_id','hadm_id')) %>%
  replace_na(list(transplant = 0)); dim(out_tb)

# other calculations
out_tb <- out_tb %>%
  mutate(
    preiculos = round(as.numeric(difftime(intime, admittime, units='days')), digits=3),
    iculos = round(as.numeric(difftime(outtime, intime, units='days')), digits=3), 
    bmi = round(weight / (height/100)^2, digits=1)
  )
```

```{r read_omr}
`physionet-data.mimic_hosp.omr` <- read.csv(file='../../hosp/omr.csv.gz', nrows=row_limit)
```

```{r check_bmi}
tmp <- `physionet-data.mimic_hosp.omr` %>%
  filter(result_name == 'BMI (kg/m2)') %>% 
  mutate(
    chartdate = as.Date(chartdate),
    bmi = as.numeric(result_value)
  ) %>%
  filter(10 <= bmi & bmi < 100) %>%
  dplyr::select(
    subject_id, chartdate, bmi
  )

tmp <- tmp %>%
  inner_join(
    dplyr::select(out_tb, subject_id, intime), 'subject_id', relationship='many-to-many'
  ) %>%
  mutate(
    indate = as.Date(intime),
    dif = abs(as.numeric(indate) - as.numeric(chartdate)),
    prior_flag = as.numeric(chartdate <= indate)
  )

tmp <- tmp %>%
  arrange(subject_id, intime, prior_flag, dif) %>%
  group_by(subject_id, intime, prior_flag) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  dplyr::select(
    subject_id, intime, prior_flag, dif, bmi 
  )

out_tb <- out_tb %>%
  left_join(filter(tmp, prior_flag == 1), c('subject_id','intime')) %>%
  dplyr::select(
    -prior_flag
  )

out_tb <- out_tb %>%
  left_join(filter(tmp, prior_flag == 0), c('subject_id','intime')) %>%
  dplyr::select(
    -prior_flag
  )

out_tb <- out_tb %>%
  mutate(
    bmi = coalesce(bmi.x, bmi.y, bmi)
  ) %>%
  dplyr::select(
    -ends_with(".x"), -ends_with(".y")
  )
```

```{r check_height}
tmp <- `physionet-data.mimic_hosp.omr` %>%
  filter(result_name == 'Height (Inches)') %>%
  mutate(
    chartdate = as.Date(chartdate),
    height = as.numeric(result_value)*2.54
  ) %>%
  filter(100 <= height & height < 250) %>%
  dplyr::select(
    subject_id, chartdate, height
  )

tmp <- tmp %>%
  inner_join(
    dplyr::select(out_tb, subject_id, intime), 'subject_id', relationship='many-to-many'
  ) %>%
  mutate(
    indate = as.Date(intime),
    dif = abs(as.numeric(indate) - as.numeric(chartdate)),
    prior_flag = as.numeric(chartdate <= indate)
  )

tmp <- tmp %>%
  arrange(subject_id, intime, prior_flag, dif) %>%
  group_by(subject_id, intime, prior_flag) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  dplyr::select(
    subject_id, intime, prior_flag, dif, height 
  )

out_tb <- out_tb %>%
  left_join(filter(tmp, prior_flag == 1), c('subject_id','intime')) %>%
  dplyr::select(
    -prior_flag
  )

out_tb <- out_tb %>%
  left_join(filter(tmp, prior_flag == 0), c('subject_id','intime')) %>%
  dplyr::select(
    -prior_flag
  )

out_tb <- out_tb %>%
  mutate(
    height = coalesce(height.x, height.y, height)
  )

out_tb <- out_tb %>%
  mutate(
    bmi = if_else(is.na(bmi), weight / (height/100)^2, bmi)
  ) %>%
  dplyr::select(
    -ends_with(".x"), -ends_with(".y")
  )
```

```{r pneumonia_datetime}
# pneumonia
tmp <- `physionet-data.mimic_hosp.microbiologyevents` %>%
  filter(spec_type_desc == 'SPUTUM') %>%
  filter(org_name != '' & org_itemid != 90856) %>%
  mutate(
    charttime = as.POSIXct(charttime, format="%Y-%m-%d %H:%M:%S"), 
    positiveculture = 1
  ) %>%
  dplyr::select(
    subject_id, charttime, positiveculture
  ) %>%
  distinct()

tmp <- tmp %>%
  inner_join(dplyr::select(
    out_tb, subject_id, hadm_id, stay_id, admittime, dischtime), 'subject_id', relationship='many-to-many'
  )

tmp <- tmp %>%
  filter(as.Date(admittime) <= as.Date(charttime) & as.Date(charttime) <= as.Date(dischtime)) %>%
  arrange(subject_id, hadm_id, stay_id, charttime) %>%
  group_by(subject_id, hadm_id, stay_id) %>%
  slice_head(n=1) %>%
  ungroup() %>%
  rename(
    pneumoniadttm = charttime
  ) %>%
  dplyr::select(
    subject_id, hadm_id, stay_id, pneumoniadttm
  )
  
out_tb <- out_tb %>%
  left_join(tmp, c('subject_id','hadm_id', 'stay_id'))
```

```{r longitudinal_vitals}
# one row for each time any vital sign was measured
vitals <- dplyr::select(`physionet-data.mimic_derived.vitalsign`, subject_id, charttime) %>%
  bind_rows(dplyr::select(`physionet-data.mimic_derived.bg`, subject_id, charttime)) %>%
  bind_rows(dplyr::select(`physionet-data.mimic_derived.ventilator_setting`, subject_id, charttime)) %>%
  distinct(); dim(vitals)

vitals <- vitals %>%
  left_join(dplyr::select(
    `physionet-data.mimic_derived.vitalsign`, subject_id, charttime, mbp, mbp_ni, heart_rate, spo2, temperature, resp_rate), 
    c('subject_id','charttime')
  ); dim(vitals)

vitals <- vitals %>%
  left_join(dplyr::select(
    `physionet-data.mimic_derived.bg`, subject_id, charttime, fio2),
    c('subject_id','charttime')
  ); dim(vitals)

vitals <- vitals %>%
  left_join(dplyr::select(
    `physionet-data.mimic_derived.ventilator_setting`, subject_id, charttime, peep, tidal_volume_set, tidal_volume_observed, tidal_volume_spontaneous, fio2),
    c('subject_id','charttime')
  ); dim(vitals)

# missing vtmode, oxdev
vitals <- vitals %>%
  mutate(
    tidal_vol = coalesce(tidal_volume_set, tidal_volume_observed, tidal_volume_spontaneous),
    map = coalesce(mbp, mbp_ni), 
    fio2 = coalesce(fio2.x, fio2.y)
  ) %>%
  rename(
    hr = heart_rate, temp = temperature, rr = resp_rate
  ) %>%
  dplyr::select(
    subject_id, charttime, peep, tidal_vol, map, hr, fio2, spo2, temp, rr
  ); dim(vitals)
```

```{r longitudinal_labs}
# labs
labs <- dplyr::select(`physionet-data.mimic_derived.complete_blood_count`, subject_id, hadm_id, charttime) %>%
  bind_rows(dplyr::select(`physionet-data.mimic_derived.chemistry`, subject_id, hadm_id, charttime)) %>%
  bind_rows(dplyr::select(`physionet-data.mimic_derived.coagulation`, subject_id, hadm_id, charttime)) %>%
  bind_rows(dplyr::select(`physionet-data.mimic_derived.bg`, subject_id, hadm_id, charttime)) %>%
  bind_rows(dplyr::select(`physionet-data.mimic_derived.enzyme`, subject_id, hadm_id, charttime)) %>%
  distinct(); dim(labs)

labs <- labs %>%
  left_join(distinct(dplyr::select(
    `physionet-data.mimic_derived.complete_blood_count`, subject_id, hadm_id, charttime, hemoglobin, platelet, hematocrit, wbc)),
    c('subject_id','hadm_id','charttime')
  ); dim(labs)

labs <- labs %>%
  left_join(distinct(dplyr::select(
    `physionet-data.mimic_derived.chemistry`, subject_id, hadm_id, charttime, glucose, potassium, sodium, creatinine, bun, bicarbonate, aniongap)),
    c('subject_id','hadm_id','charttime')
  ); dim(labs)

labs <- labs %>%
  left_join(distinct(dplyr::select(
    `physionet-data.mimic_derived.coagulation`, subject_id, hadm_id, charttime, inr, ptt)),
    c('subject_id','hadm_id','charttime')
  ); dim(labs)

labs <- labs %>%
  left_join(distinct(dplyr::select(
    `physionet-data.mimic_derived.bg`, subject_id, hadm_id, charttime, ph, pco2, po2, lactate)),
    c('subject_id','hadm_id','charttime')
  ); dim(labs)

labs <- labs %>%
  left_join(distinct(dplyr::select(
    `physionet-data.mimic_derived.enzyme`, subject_id, hadm_id, charttime, ast)),
    c('subject_id','hadm_id','charttime')
  ); dim(labs)

labs <- labs %>%
  rename(
    hgb = hemoglobin, hct = hematocrit,  hco3 = bicarbonate, paco2 = pco2, pao2 = po2, 
    anion_gap = aniongap
  ) %>%
  dplyr::select(
    subject_id, hadm_id, charttime, glucose, hgb, potassium, sodium, platelet, hct, ph, creatinine,
    wbc, bun, hco3, paco2, pao2, lactate, inr, anion_gap, ptt, ast
  ); dim(labs)
```

```{r longitudinal_vasopressors}
out_vasos <- mutate(`physionet-data.mimic_derived.dobutamine`, med = 'dobutamine') %>%
  bind_rows(mutate(`physionet-data.mimic_derived.norepinephrine`, med = 'norepinephrine')) %>%
  bind_rows(mutate(`physionet-data.mimic_derived.epinephrine`, med = 'epinephrine')) %>%
  bind_rows(mutate(`physionet-data.mimic_derived.vasopressin`, med = 'vasopressin'))
```

```{r read_d_items}
`physionet-data.mimic_icu.d_items` <- read.csv(file='../../icu/d_items.csv.gz', nrows=row_limit)
```

```{r cpn}
cpn <- `physionet-data.mimic_icu.d_items` %>%
  filter(category == 'Nutrition - Parenteral') %>%
  dplyr::select(
    itemid
  ) %>%
  distinct() %>%
  inner_join(`physionet-data.mimic_icu.inputevents`, 'itemid') %>%
  dplyr::select(
    subject_id, starttime, endtime
  )
```


```{r ringers_lactate}
ringers_lactate <- `physionet-data.mimic_icu.inputevents` %>% 
  filter(itemid %in% c(225827, 225828)) %>%
  dplyr::select(
    subject_id, starttime, endtime, amount, totalamount, ordercategoryname, ordercategorydescription
  )
```

```{r write_cohort_lab_vitals_data}
# demo data
write.csv(out_tb, file='./derived/v4/demo.csv', row.names=FALSE, na='')

# labs data
write.csv(labs, file='./derived/v4/labs.csv', row.names=FALSE, na='')

# vitals data
write.csv(vitals, file='./derived/v4/vitals.csv', row.names=FALSE, na='')

# vasoactives data
write.csv(out_vasos, file='./derived/v4/vasoactives.csv', row.names=FALSE, na='')

# ventilation
write.csv(`physionet-data.mimic_derived.ventilation`, file='./derived/v4/ventilation.csv', row.names=FALSE, na='')

# gcs
write.csv(`physionet-data.mimic_derived.gcs`, file='./derived/v4/gcs.csv', row.names=FALSE, na='')

# suspicion of infection
write.csv(`physionet-data.mimic_derived.suspicion_of_infection`, file='./derived/v4/suspicion_of_infection.csv', row.names=FALSE, na='')

# urine output rate
write.csv(`physionet-data.mimic_derived.urine_output_rate`, file='./derived/v4/urine_output_rate.csv', row.names=FALSE, na='')

# urine output
write.csv(`physionet-data.mimic_derived.urine_output`, file='./derived/v4/urine_output.csv', row.names=FALSE, na='')

# icu stay times
write.csv(`physionet-data.mimic_derived.icustay_times`, file='./derived/v4/icustay_times.csv', row.names=FALSE, na='')

# aps
write.csv(`physionet-data.mimic_derived.apsiii`, file='./derived/v4/apsiii.csv', row.names=FALSE, na='')

# sofa
write.csv(`physionet-data.mimic_derived.sofa`, file='./derived/v4/sofa.csv', row.names=FALSE, na='')

# sepsis
write.csv(`physionet-data.mimic_derived.sepsis3`, file='./derived/v4/sepsis.csv', row.names=FALSE, na='')

# saps
write.csv(`physionet-data.mimic_derived.sapsii`, file='./derived/v4/sapsii.csv', row.names=FALSE, na='')

# cpn
write.csv(cpn, file='./derived/v4/cpn.csv', row.names=FALSE, na='')

# ringers lactate
write.csv(ringers_lactate, file='./derived/v4/ringers_lactate.csv', row.names=FALSE, na='')
```

```{r}
# End
```

